---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
For data collection both in-house interviewers and contracts with telephone callcenters are used. Random Digit Dialing used for sample randomization. Interview and calls are common cause of non-response bias. Method of calling might cause bias of availability - as not all citizens have home phones. Some false information we can conduct out is filtered out using weighting and stratification variables.

* * *

## Part 2: Research questions

**Research quesion 1:**

1) Is there any correlations between Physical Activity of person and Depression? As we have idea that higher physical activity is associated with lower level of depression and vise versa.

**Research quesion 2:**

2) Are there any correlations between prediabetes and usage of sweetened and sugared drinks? Which percentage of diagnosed with prediabetes used one, both or none of these?

**Research quesion 3:**

3) Do we have any interconnection of health days and sleep time? How does those variable correlate with general health of respondent?

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
X_totinda: Leisure Time phys activity
addepev2: ever told you had a depressive disorder

```{r}
brfss2013 <- brfss2013 %>% mutate(phys_act = ifelse( X_totinda == "Had physical activity or exercise", TRUE, FALSE));
# filter n/a of the data
sub_brfss <- dplyr::filter(brfss2013, !is.na(phys_act))
sub_brfss <- dplyr::filter(sub_brfss, !is.na(addepev2))
sub_brfss %>%
  count(phys_act, addepev2)%>%
  group_by(addepev2)%>%
  mutate(pct = n/sum(n))%>%
  ggplot(aes(x = addepev2, y = pct, fill = phys_act)) +
  geom_col(width=0.7) +
  geom_text(aes(label = paste0(round(pct*100), '%')),
            position=position_stack(vjust = 0.5))
```
Yes/no - indicates if you where diagnosed with depression, and bar - percentage of physical activity
As we see there are more physicly active people in group of non-depression, and strong difference between those diagnosed with depression. We assume that there is some correlation between those variables

**Research quesion 2:**

```{r}
#filter all rows which does not have information
fl_brfss <- brfss2013 %>% filter(!is.na(ssbsugar)) %>% filter(!is.na(ssbfrut2))
#examine distribution of oftennes sugar drinks
#ggplot(fl_brfss, aes(ssbsugar)) + geom_bar()
sugar_mean = mean(fl_brfss$ssbsugar)
#ggplot(fl_brfss, aes(ssbfrut2)) + geom_bar()
frut_mean = mean(fl_brfss$ssbfrut2)
#we assume that often for sugar drinks and sweetened fruit drinks is more than 200
#let's create variables which describe frequence of sugar and sweetened drinks usage
fl_brfss <- fl_brfss %>% mutate(freq_sugar = ifelse(ssbsugar > sugar_mean, TRUE, FALSE)) %>% mutate(freq_frut = ifelse(ssbfrut2 > frut_mean, TRUE, FALSE))
#create frequence table of 2 variables
df <- as.data.frame(table(fl_brfss$freq_frut, fl_brfss$freq_sugar))
#create variable of 2 factors
df <- df %>% mutate(name = as.numeric(Var1) + as.numeric(Var2) - 1)
ggplot(df, aes(x="", y=Freq, fill=name))+
geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0)
```
Pie chart of people diagnosed with prediabetes. 1 - absense of cofactors, 2 - usage of sugar or sweetened juice, 3 - frequent usage of sugar and sweetened juices.
As we can see - large part of prediabetes diagnosed group are often using sweetened drinks and juices. We can assume that there is some trend with those variables

**Research quesion 3:**

```{r}
#sleptim1 genhlth poorhlth
#read new dataset of unique healthy days poorhlth
poor_hlth <- brfss2013 %>% filter(!is.na(sleptim1)) %>% filter(!is.na(genhlth)) %>% filter(!is.na(poorhlth)) %>% select(c(sleptim1, genhlth, poorhlth)) %>% mutate(genhlth=as.numeric(genhlth))
poor_hlth <- aggregate(poor_hlth, list(Days = poor_hlth$poorhlth), mean)
ggplot(poor_hlth, aes(y = sleptim1, x = poorhlth, fill = genhlth)) + geom_bar(stat="identity") + geom_smooth(method = "lm")
```
Poorhlth - number of days in month when respondent feeling bad, sleptim1 - hours of sleep he have on average and genlth - how assess his own health.
As we see, genhlth strongly correlates with poorhlth - more days on month respondend feels bad - worsen he feels in general. There is trend of sleep time and poor health - less sleep time - more bad health days respondent have.

